---
title: "Exercise 3"
author: "Utkarsh Pal"
date: "2023-03-28"
output: pdf_document
---

# Advice Network at USPTO

## Loading the Packages:
```{r Loading Packages}
library(readr)
library(arrow)
library(gender)
library(dplyr)
library(tidyr)
library(wru)
library(ymd)
library(lubridate)
library(stringr)
library(ggplot2)
library(igraph)
```

## Loading the files and adding gender, race, and tenure for each examiner
```{r Data Load}
applications = read_parquet("C:\\Users\\Utkarsh\\Desktop\\app_data_sample.parquet")
edges = read_csv("C:\\Users\\Utkarsh\\Desktop\\edges_sample.csv")

# Gender
examiner_names <- applications %>% distinct(examiner_name_first)
examiner_names_gender <- examiner_names %>% 
  do(results = gender(.$examiner_name_first, method = "ssa")) %>% 
  unnest(cols = c(results), keep_empty = TRUE) %>% 
  select(examiner_name_first = name, gender, proportion_female)
applications <- applications %>% 
  left_join(examiner_names_gender, by = "examiner_name_first")

# Race
examiner_surnames <- applications %>% select(surname = examiner_name_last) %>% distinct()
examiner_race <- predict_race(voter.file = examiner_surnames, surname.only = T) %>% as_tibble()
examiner_race <- examiner_race %>% 
  mutate(max_race_p = pmax(pred.asi, pred.bla, pred.his, pred.oth, pred.whi)) %>% 
  mutate(race = case_when(
    max_race_p == pred.asi ~ "Asian",
    max_race_p == pred.bla ~ "black",
    max_race_p == pred.his ~ "Hispanic",
    max_race_p == pred.oth ~ "other",
    max_race_p == pred.whi ~ "white",
    TRUE ~ NA_character_
  ))
applications <- applications %>% 
  left_join(examiner_race, by = c("examiner_name_last" = "surname"))


# Tenure
examiner_dates <- applications %>% 
  select(examiner_id, filing_date, appl_status_date)
examiner_dates <- examiner_dates %>% 
  mutate(start_date = ymd(filing_date), end_date = as_date(dmy_hms(appl_status_date)))
examiner_dates <- examiner_dates %>% 
  group_by(examiner_id) %>% 
  summarise(
    earliest_date = min(start_date, na.rm = TRUE), 
    latest_date = max(end_date, na.rm = TRUE),
    tenure_days = interval(earliest_date, latest_date) %/% days(1)
  ) %>% 
  filter(year(latest_date)<2018)
applications <- applications %>% left_join(examiner_dates, by = "examiner_id")

applications <- applications %>%
  mutate(tenure_years = tenure_days / 365) %>%
  mutate(tenure_range = case_when(
    tenure_years < 6  ~ '<6',
    tenure_years <= 10 ~ '6-10',
    tenure_years <= 15 ~ '11-15',
    tenure_years <= 19 ~ '16-19',
    tenure_years <= 100 ~ '20+',
    TRUE ~ NA_character_
  ))

# Workgroups
applications <- applications %>% 
  mutate(examiner_workgroup = str_sub(examiner_art_unit, 1, -2))
# Drop NAs
applications <- applications %>% drop_na(gender, tenure_days, race)
```

## Filtering Work Groups: 171 & 172
```{r Work Groups}
# Picking two work groups:
wrkgrp_1 <- "171"
wrkgrp_2 <- "172"

wrkgrp <- applications %>%
  filter(examiner_workgroup %in% c(wrkgrp_1, wrkgrp_2))
```

## Summary Statistics & Plots
### Gender
```{r Gender}
gender <- wrkgrp %>%
  group_by(examiner_workgroup, gender) %>%
  summarise(count = n()) %>%
  group_by(examiner_workgroup) %>%
  mutate(total_count = sum(count),
         percentage = count/total_count * 100) %>%
  select(-total_count) %>%
  arrange(examiner_workgroup, count, .by_group = TRUE)

gender_plot <- wrkgrp %>%
  ggplot(aes(x = gender, fill = gender)) + geom_bar() +
  facet_wrap(~examiner_workgroup) +
  labs(title = "Gender Distribution",
       x = "Gender",
       y = "Count") +
  theme(axis.text.x = element_blank())

### Summary Statistics
print(gender)
### Plot
print(gender_plot)
```

### Race
```{r Race}
race <- wrkgrp %>%
  group_by(examiner_workgroup, race) %>%
  summarise(count = n()) %>%
  group_by(examiner_workgroup) %>%
  mutate(total_count = sum(count),
         percentage = count/total_count * 100) %>%
  select(-total_count) %>%
  arrange(examiner_workgroup, count, .by_group = TRUE)

race_plot <- wrkgrp %>%
  ggplot(aes(x = race, fill = race)) + geom_bar() +
  facet_wrap(~examiner_workgroup) +
  labs(title = "Race Distribution",
       x = "Race",
       y = "Count") +
  theme(axis.text.x = element_blank())

### Summary Statistics
print(race)
### Plot
print(race_plot)
```


### Tenure (Years)
```{r Tenure}
tenure <- wrkgrp %>%
  group_by(examiner_workgroup, tenure_range) %>%
  summarise(count = n()) %>%
  group_by(examiner_workgroup) %>%
  mutate(total_count = sum(count),
         percentage = count/total_count * 100) %>%
  select(-total_count) %>%
  arrange(examiner_workgroup, count, .by_group = TRUE)

tenure_plot <- wrkgrp %>%
  ggplot(aes(x = tenure_range, fill = tenure_range)) + geom_bar() +
  facet_wrap(~examiner_workgroup) +
  labs(title = "Tenure Distribution",
       x = "Tenure Range",
       y = "Count") +
  theme(axis.text.x = element_blank())

### Summary Statistics
print(tenure)
### Plot
print(tenure_plot)
```

### Work groups 171 and 172 are very similar in terms of demographic attributes after removing observations with missing values. There are a majority of male examiners at 67.87% in work group 171 compared to female examiners at 32.13%. Similarly, in work group 172, 71.56% of examiners are male and the remaining 28.44% are female. Both work groups have a majority of White examiners, a small yet significant minority of Asians, and a very small minority of Black and Hispanic examiners. 74.78% of examiners in work group 171 are White, 19.15% are Asian, 3.01% are Hispanic, and 3.04% are Black. In work group 172, the breakdown is 82.65% White, 13.82% Asian, 2.48% Hispanic, and 1.04% Black. Both work groups have a large number of examiners with a high level of tenure. In both work groups, the overwhelming majority (greater than 60%) of examiners fall under 16-19 years of tenure. The next largest cluster for both work groups is the 11-15 years of tenure (both have over 25% in this cluster).

## Creating advice networks from 'edges_sample'
### I will be using degree and betweenness centrality, since the former is helpful in identifying the most "popular" or "well-connected" nodes in the network and the latter is useful for identifying the nodes that are most critical for facilitating communication and coordination between different parts of the network.

### Plot
```{r Plot2}
g <- graph_from_data_frame(edges[, c("ego_examiner_id", "alter_examiner_id")], directed = TRUE)
matcher <- applications %>% 
  select(examiner_id, examiner_workgroup) %>% 
  distinct()

V(g)$workgroup <- matcher$examiner_workgroup[match(V(g)$name, matcher$examiner_id)]
gNew <- g %>% 
  induced_subgraph(V(g)[V(g)$workgroup %in% c(wrkgrp_1, wrkgrp_2)])

colors_condition <- ifelse(V(gNew)$workgroup == wrkgrp_1, 'blue', 'red')
plot(gNew, vertex.color = colors_condition,
     vertex.label = NA,
     vertex.size = 7,
     edge.arrow.size = 0.3,
     main = "Work Groups 171 (Blue) and 172 (Red)")
```

### Centralities
```{r Centralities}
degree_centrality <- degree(gNew, mode = "all")
betweenness_centrality <- betweenness(gNew, directed = TRUE)
V(gNew)$degree_centrality <- degree_centrality
V(gNew)$betweenness_centrality <- betweenness_centrality
centralities <- data.frame(
  examiner_id = V(gNew)$name,
  workgroup = V(gNew)$workgroup,
  degree_centrality = V(gNew)$degree_centrality,
  betweenness_centrality = V(gNew)$betweenness_centrality
)
class(centralities$examiner_id)
centralities$examiner_id <- as.numeric(centralities$examiner_id)

connections <- applications %>%
  select(examiner_id, gender, race, tenure_days) %>%
  inner_join(centralities, by = "examiner_id") %>%
  distinct(examiner_id, .keep_all = TRUE)
print(connections)
```

